Toward a clearer understanding of the inverse association between cancer and dementia

### 2021-04-28 06_presentations/ser2021/index.Rmd
presentation almost done

@@

.right[L. Paloma Rojas-Saunero, MD

**Epidemiology department**]

@@

---
## Application to the Rotterdam Study

@@

- The process of connecting our real question to the data help us identify the structure that will later represent the analytic decisions and assumptions.

@@

- Cancer as the proxy for a molecular mechanism is tricky...
--

- When death is a competing event, prior to choosing the estimator we need to define what is the estimand of interest and be explicit about the assumptions and interpretations.

>>>>

<br>

@@

.right[

L. Paloma Rojas-Saunero, MD

**Epidemiology department**

`r icons::fontawesome$brands$twitter` @palolili23

@@

---
### Application to the Rotterdam Study

@@

- Defining the underlying mechanism behind cancer diagnosis helps in multiple way.

  + Helps identify sources of bias and guides analysis

  + Opens conversations about future data collection

@@

- Since death acts as a competing event, we first need to define the estimand of interest.

  + The direct effect relies on the independent censoring assumption

  + Interpretation implies preventing death.

  + The separable effects provides a clearer intuition to address this question.

---
class: middle

### Acknowledgments

- Sonja A. Swanson

- Kimberly van der Willik

- Arfan M. Ikram

---
class:middle

### Gracias!   Thank you!

`r icons::fontawesome("paper-plane", style = "solid")` </i>&nbsp; l.rojassaunero@erasmusmc.nl </a> <br>

`r icons::fontawesome$brands$twitter` @palolili23

`r icons::fontawesome$brands$github` @palolili23
